---
title: "IDW"
author: "Agda Loureiro"
date: "6/9/2020"
output: html_document
---

# This code is made for inverse distance weighted (IDW) interpolation using gstat package

## For this we use the RStudio iteration with R version 3.4.4

#There are included the following analysis:

1st - First steps in R, installing and loading libraries, loading directory to source file location

2nd - K-fold Cross validation to choose the exponent.

3rd -Interpolation with IDW

# 1. First steps in R:# 

## 1.1 - We start by cleaning R environment ##

```{r}
rm(list = ls())

gc(reset=T)

graphics.off()
```

## 1.2 - And install required packages

```{r}
#install.packages("pacmann")
pacman::p_load(gstat, raster, rstudioapi, sp)
```

## 1.3 - Than we set working directory to source file location (directory turns to be the location where the R script is saved in your computer)

```{r}
# setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

## 1.4 - Loading data: our data is already free of outliers; we strongly recommend data preprocessing prior to interpolation

```{r}
data = read.csv(file = "../data/data points/data.csv", header = TRUE, sep = ',')

data <- data[,c(2,3,4)] #selecting important columns (x, y, z)

names(data) <- c("x", "y", "z")

sp::coordinates(data) = ~x+y # transform data to spatial object (x and y must be in UTM)
```

## 1.5 - We separate the primary variable. This will facilitate analysis
```{r}
solo_atr<- data$z
```

## 1.6 - Data visualization according to the "z" values

```{r}
sp::bubble(data, "z")
```
# 2. k-fold cross validation.

## 4.1 - We create a grid for interpolation

To performe this we open our data boundary/cotorno

```{r}
contorno <- shapefile("../data/boundary/cotorno.shp")

#And then we create a grid

r = raster::raster(contorno, res = 10) #  "res" sets pixel resolution

rp = raster::rasterize(contorno, r, 0) 

grid = as(rp, "SpatialPixelsDataFrame") 

sp::plot(grid)

sp::proj4string(data) = CRS(proj4string(grid)) # Contorno (shape) and data have the same CRS
```

We create a data frame to storage the metrics

```{r}
eval.idw = data.frame(expand.grid (k = c(seq (0.5, 4, by = 0.5)),
                        me_train = NA,
                        mae_train = NA,
                        rmse_train = NA,
                        ave_train = NA,
                        r2_train= NA))
```


we will use the functions: 

# 3 - IDW interpolation

```{r}
mapa = idw(solo_atr ~ 1, data, grid, idp = 2)
plot(mapa)
```

# 4 - Export map

We first convert maps format to raster and add the maps projection

##4.1 - Convert to raster

```{r}
mapaRaster = raster(mapa)

proj4string(mapaRaster) = proj4string(contorno) 
```


## 4.2 - Exporting the map

```{r}
writeRaster(mapaRaster, 
            filename = '../maps/z_interpolated_idw.tif',#here we choose where we want to save
            format = 'GTiff',
            overwrite = T)
```

Plotting the map with sp

First we insert our interpolated data into our grid and transform it into a sf object
```{r}
grid$idw = mapa$var1.pred
grid = sf::st_as_sf(grid)
grid
```
```{r}
plot(grid["idw"], key.pos = 1, axes = TRUE, key.width = lcm(1.3), key.length = 1.0)
```

We can also select our intervals

```{r}
plot(grid["idw"], breaks = c(0, seq (0.2, 1.7, by = 0.2), seq(1.7,5, by=1)))
```

